---
title: "Solutions to Exam Questions Pig Science Breeding"
author: "Peter von Rohr"
date: "`r Sys.Date()`"
output:
  word_document:
    reference_docx: docx/exam_questions_template.docx
  pdf_document:
    includes:
      in_header: tex/header.tex
  html_document:
    df_print: paged
---

\newcommand{\points}[1]
{\begin{flushright}\textbf{#1}\end{flushright}}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# rmdhelp::show_knit_hook_call()
knitr::knit_hooks$set(hook_convert_odg = rmdhelp::hook_convert_odg)
```

```{r points}
n_points_q1 <- 6
n_points_q2 <- 4
n_points_q3 <- 4
n_points_q4 <- 3
n_points_q5 <- 3
n_points_q6 <- 6
n_points_total <- n_points_q1 + n_points_q2 + n_points_q3 + n_points_q4 + n_points_q5
tbl_points <- tibble::tibble(Question = c(as.character(1:5), "Total"),
                             `Maximum Points` = c(n_points_q1,
                                                  n_points_q2,
                                                  n_points_q3,
                                                  n_points_q4,
                                                  n_points_q5,
                                                  n_points_total))
knitr::kable(tbl_points,
             booktabs = TRUE,
             longtable = TRUE)
```


# Question 1
What are the six components of a breeding program?
  
_Wie lauten die sechs Komponenten eines Zuchtprogramms?_

```{r, echo=FALSE, results='asis'}
cat("__", n_points_q1, "__", "\n", sep = "")
```



# Question 2
Compare a typical breeding program in cattle with a breeding program in pigs. Draw the characteristic scheme for the two types of breeding programs

_Vergleichen sie ein typisches Zuchtprogramm beim Rind mit einem typischen Zuchtprogramm beim Schwein. Zeichnen sie dabei die charakteristischen Schemata der beiden Zuchtprogrammtypen auf._

```{r, echo=FALSE, results='asis'}
cat("__", n_points_q2, "__", "\n", sep = "")
```



# Question 3

What is the goal of model selection and what are two well-known approximations to do model selection?

_Was ist das Ziel der Modellselektion und wie lauten zwei bekannte Approximationen um Modellselektion auszuführen?

```{r, echo=FALSE, results='asis'}
cat("__", n_points_q3, "__", "\n", sep = "")
```



# Question 4
Why are variance components important in the process of a genetic evaluation? Name two methods how to estimate variance components.

_Weshalb sind Varianzkomponenten wichtig im Prozess der genetischen Auswertung? Benennen sie zwei Methoden um Varianzkomponenten zu schäten._

```{r, echo=FALSE, results='asis'}
cat("__", n_points_q4, "__", "\n", sep = "")
```



# Question 5
What are three different models to predict breeding values?

_Wie lauten drei verschiedene Modelle zur Schätzung von Zuchtwerten?_


```{r, echo=FALSE, results='asis'}
cat("__", n_points_q5, "__", "\n", sep = "")
```  



# Questions 6
What are the four components that determine the selection response per year? Which of the two components are antagonistic? What is the recently introduced breeding method that resolves the antagonism?

_Wie lauten die vier Komponenten, welche den Zuchtfortschritt pro Jahr bestimmen? Welche der beiden Komponenten stehen antagonistisch zueinander? Welche kürzlich eingeführte Züchtungsmethode kann diesen Antagonismus auflösen?_

```{r, echo=FALSE, results='asis'}
cat("__", n_points_q5, "__", "\n", sep = "")
```  

